### 2025-03-31 writing/appendix-1.Rmd

---
title: \large A new global measure of environmental unpredictability 
subtitle: Appendix 1
author:
  - name: Stefano Mezzini
    institute: [braes, biol]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
    correspondence: "yes"
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'ndvi-stochasticity.bib'
csl: 'freshwater-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font={scriptsize, doublespacing}, labelfont=bf, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
# https://github.com/pandoc/lua-filters/blob/master/author-info-blocks/README.md
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: false
    keep_tex: false # for submitting TeX file
editor_options:
  chunk_output_type: console
---

\clearpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%')
```

```{r packages, warning=FALSE}
library('knitr')
```

\noindent The figures below illustrate some of the data used for fitting the models. All color palettes below are available via the `khroma` package [@frerebeau_khroma_2024] for `R` [@r_core_team_r_2024]. The code for generating the figures is available [on GitHub](https://github.com/QuantitativeEcologyLab/ndvi-stochasticity/blob/main/analysis/figures/input-data.R).

```{r ecoregions, fig.cap="Map of the ecoregions."}
include_graphics('../figures/input-data/ecoregions.png')
```

```{r polygons, fig.cap="Map of the polygons for each ecoregion. Each polygon is colored with one of 9 colors based on the polygon names. Note that some neighboring polygons have the same color."}
include_graphics('../figures/input-data/polygons.png')
```

```{r elevation, fig.cap="Map of elevation above sea level."}
include_graphics('../figures/input-data/elev-m.png')
```

```{r distance-coast, fig.cap="Map of the estimated distance from the nearest coastal edge."}
include_graphics('../figures/input-data/dist-m.png')
```

```{r n-rasters, fig.cap="Number of days with a raster within 15-day periods starting on January $1^{\\text{st}}$, for each year."}
include_graphics('../figures/input-data/n-rasters-time.png')
```

\clearpage

# References