### 2025-04-14 writing/manuscript.Rmd
started writing methods

---
title: "DENVar: A global Dynamic Estimate of NDVI Variance"
author:
  - name: Stefano Mezzini
    institute: [braes, biol]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
    correspondence: "yes"
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'ndvi-stochasticity.bib'
csl: 'freshwater-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font={scriptsize, doublespacing}, labelfont=bf, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
# https://github.com/pandoc/lua-filters/blob/master/author-info-blocks/README.md
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: false
    keep_tex: false # for submitting TeX file
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE}
# need to write abstract before cover page to state number of words in it
abstract <- "Since the 1980s, remote-sensing estimates of habitat productivity such as the Normalized Differnece Vegetation Index (NDVI) have helped develop and support a variety of hypotheses and inform policy and industriy-related decisions. The free availability of such data has allowed widespread access and use of well-established metrics of habitat productivity. However, historically, users tended to focus solely on (predictable) trends in productivity. In recent years, people have been recognizing the importance of envrironmental unpredictability (stochasticity), particularly as climate change and human-induced rapid environmental change transform the ecosystems species have evolved in and adapted to. We adress this gap by presenting a new global Dynamic Estimate of NDVI Variance, *DENVar*. We estimate spatiotemporal trends in mean NDVI and the variance around the mean using Hierarchical Generalized Additive Models, which provide a flexible yet transparent model structure, unlike machine learning approaches such as neural netwoks. We show that DENVar can be used as a reliable proxy of environmental stochasticity, strongly correlates with ***XXX***, and can be used to test many hepotheses related to environmental stochasitcity, such as forage unpredictability, regime shifts, ... . We conclude by offering some considerations around quantifying and interpreting envirnomental stochasticity, and we provide a publicly and freely available shiny app for estimating mean NDVI and the variance around it, DENVar, for given coordinates and optional dates. The app requires no knowledge of statistics or coding. Additionally, we provide a temporally static raster of mean NDVI and DENVar for the years 1981-2025 that can be used for GIS applications and does not require any knowledge of $\\texttt{R}$."

if(FALSE) {
  stringi::stri_count_words(abstract)
}
```

<!-- elements of the title page -->

\clearpage

\noindent \textbf{Article type}: Research article

\noindent \textbf{Words in abstract}: `r stringi::stri_count_words(abstract)`

\noindent \textbf{Words in main text}: `r suppressMessages(wordcountaddin::word_count())`

\noindent \textbf{Figures}: \textcolor{red}{\textbf{XXX}}

\noindent \textbf{Tables}: \textcolor{red}{\textbf{XXX}}

\noindent \textbf{References}: \textcolor{red}{\textbf{XXX}}

\noindent \textbf{Appendices}: \textcolor{red}{\textbf{XXX}}

\noindent \textbf{Key words:} remote sensing, NDVI, environmental stochasticity, \textcolor{red}{\textbf{XXX}}

<!-- start main body on a new page -->

\newpage

\doublespacing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%', fig.align = 'center')
```

\linenumbers

# Abstract

\noindent `r abstract`

\newpage

# Introduction

<!-- remote sensing allows us to study landscapes in detail and over large spatiotemporal scales -->

\noindent Developments in remote sensing since the 1970s have greatly expanded our ability to quantify and monitor landscapes using satellite imagery

- NDVI is an estimate of greenness, NPP, forage availability
- NDVI has many uses and is often convenient because it's free, well-established, and

<!-- successes of NDVI: pettorelli, Aikens -->

many people focus on changes in the mean

<!-- what counts as a change? predictable vs stochastic changes -->

ability to predict and respond to changes in mean conditions (forage availability, phenology, etc.) depends on the scale of changes relative to the perception scale [@levin_problem_1992; @frankham_importance_2004; @riotte-lambert_environmental_2020; @steixner-kumar_strategies_2020] and memory [@abrahms_memory_2019]. Consequently, environmental stochasticity is also an important driver, besides changes in the mean [@mezzini_how_2025]

<!-- difficulties in estimating variance -->

* variance is hard to calculate, as it depends on a correct estimate of the mean
* paper on estimating the asymptote of environmental variance globally, but variance for a closed system must converge towards a finite value [@bachmaier_variogram_2011; @fleming_fine-scale_2014]

<!-- we present a new measure that is estimated in an understandable way -->

* we present DENVar
* HGAMs are higly flexible models that allow us to estimate trends in the data without imposing preconceived expectations
* more transparent than black-box machine learning methods like neural networks, etc.
* https://www.sciencedirect.com/science/article/abs/pii/S0034425718305625
* https://silvis.forest.wisc.edu/data/dhis/
* we provide global rasters of the mean and variance in NDVI, averaged across 1981-2025
* we provide a shiny app for people to calculate mean and variance without any knowledge of `R`

# Methods

## Choice of color schemes

We represent NDVI using a modified version of Fabio Crameri's divergent *bukavu* palette [@crameri_geodynamic_2018], which has high-contrast for deuteranope and protanope vision. For NDVI values between 0 and 1 (the right half of the palette), the colors also have sufficient contrast for the colors to be distinguishable by both tritanope and achomatic vision. Appendix A contains an approximate representation of the color palette for each vision type. We obtained all color palettes from the `khroma` package [v. 1.14.0, @frerebeau_khroma_2024] for `R` [v. 4.4.1, @r_core_team_r_2024].

## Input data

We obtained NDVI data using the image composites from the AVHRR and VIIRS sensors [@vermote_noaa_2018; @vermote_noaa_2022]. The code for downloading the data directly from the NOAA server is available at [https://github.com/QuantitativeEcologyLab/ndvi-stochasticity/blob/main/analysis/002-download-ndvi-rasters.R](https://github.com/QuantitativeEcologyLab/ndvi-stochasticity/blob/main/analysis/002-download-ndvi-rasters.R). The number of non-NA land raster cells was substantially larger than the maximum data frame size in `R` ($2.28 \times 10 ^{10} > 2^{31} - 1 \approx 0.21 \times 10^{10}$), so we reduced the dataset size by calculating 15-day averages of NDVI (see Fig. A2) and aggregating the averaged rasters by a factor of $2 \times 2$ with `terra::aggregate(fact = 2)`. Although this resulted in a reduction in sample size, the temporal averaging also reduced signal-related noise.

The predictor data for the models included: WWF biome [@olson_global_2002, see Fig. A3], ecoregion [the individual polygons from @olson_global_2002, see Fig. A4], integer day of year (1 to 366), integer year (1981 to 2025), and elevation above sea level (Fig. A5). We excluded all polygons that did not have any NDVI data (0.06% of land, 29% of the polygons, maximum area: $452~\text{km}^2 \approx (21~\text{km})^2$). We distinguished between biomes in the northern and southern hemisphere to allow for different seasonal trends between hemispheres. The 15-day averages allowed us to have non-repeating values of day of year across years, since 15 is not a factor of 365 or 366. We downloaded the global digital elevation model using the `get_elev_raster()` function from the `elevatr` package for `R` [v. 0.99.0, @hollister_elevatr_2023] with a resolution of 0.076 degrees.

```{r, eval=FALSE}
# approximate estimate of the number of unique DOY values
length(unique(lubridate::yday(as.Date('1981-09-01') + seq(0, 15800, by = 15))))
```

## Modeling

We estimated spatiotemporal trends in mean NDVI using Hierarchical Generalized Additive Models (HGAMs) via the `mgcv` package for `R` [v. 1.9-3, @wood_generalized_2017]. To reduce modeling fitting times with negligible losses to model accuracy, we used the `bam()` function with fast REstricted Marginal Likelihood (`method = fREML`) and covariate discretization (`discrete = TRUE`). See @wood_generalized_2015 and @wood_generalized_2017-1 for more information. Ideally, NDVI should be modeled using beta location-scale models (after the linear transformation $Y^*=\frac{Y + 1}{2}$) to account for the fact that: (1) NDVI is bounded between -1 and 1, and (2) the variance in NDVI is dependent on the mean (and vice-versa), since ecosystems with very low NDVI (e.g., rock, ice, or concrete) or very high NDVI (e.g., dense forest) tend to have lower variance in NDVI. However, preliminary tests showed that fitting times for beta models were prohibitive, especially for beta location-scale models. Additionally, Gaussian models fit substantially faster and provided very similar spatialtemporal estimates of mean NDVI (see the `analysis/000-sardinia-test.R` script in the GitHub repository). The model for the mean NDVI had the structure below:

\singlespacing

\scriptsize

```{r m_mean, eval=FALSE, echo=TRUE}
m_mean <- bam(
  ndvi_15_day_mean ~
    biome + # to avoid intercept shrinkage
    s(poly_id, bs = 'mrf', xt = list(nb = nbs)) +
    s(doy, biome, bs = 'fs', xt = list(bs = 'cc'), k = 10) +
    s(year, biome, bs = 'fs', xt = list(bs = 'cr'), k = 10) +
    ti(doy, year, biome, bs = c('cc', 'cr', 're'), k = c(5, 5)) +
    s(elevation_m, bs = 'cr', k = 5),
  family = gaussian(),
  data = d,
  method = 'fREML',
  knots = list(doy = c(0.5, 366.5)),
  drop.unused.levels = TRUE,
  discrete = TRUE,
  samfrac = 0.001, # find initial guesses with a subset of the data
  nthreads = future::availableCores(logical = FALSE) - 2,
  control = gam.control(trace = TRUE))
```

\normalsize

\doublespacing

Biome-specific fixed-effect intercepts allowed to account for the differences in mean NDVI throught biomes without coefficient shrinkage (i.e., reversion towards the global mean). A Markov Random Field of the ecoregions [`bs = mrf`, see page ***240?*** of @wood_generalized_2017] estimated smoothed, ecoregion-specific deviations from the biome-level spatial means. Two factor-smooth interaction terms (`bs = 'fs'`) accounted for biome-specific seasonal and yearly trends. Seasonal terms were made cyclical by using cyclical cubic splines (`xt = list(bs = 'cc')`), which are continuous up to and including second derivative at the edge knots (0.5 and 366.5, which correspond to 00:00 of January 1^st^ and 24:00 of December 31^st^, respectively). A tensor product interaction smooth for each biome estimated the change in seasonal trends over the years. Finally, a smooth of elevation above sea level accounted for the effect of altitude. See @pedersen_hierarchical_2019 for more information on hierarchical modeling and the use and interpretation of factor smooth interaction terms.

***will we subtract the average residual from each pixel?***

A second HGAM estimated the variance in NDVI around the mean estimated by the model above. The model had an identical structure to the model for the mean, with the exception that (1) the response variable was the squared residual from the first model, such that the model estimated the mean squared residual (i.e., the variance) for a given point in time and space, and (2) a smooth term of the estimated mean NDVI. The model is available below:

\singlespacing

\scriptsize

```{r m_var, eval=FALSE, echo=TRUE}
m_var <- bam(
  e_2 ~
    biome +
    s(poly_id, bs = 'mrf', xt = list(nb = nbs)) +
    s(doy, biome, bs = 'fs', xt = list(bs = 'cc'), k = 10) +
    s(year, biome, bs = 'fs', xt = list(bs = 'cr'), k = 10) +
    ti(doy, year, biome, bs = c('cc', 'cr', 're'), k = c(5, 5)) +
    s(elevation_m, bs = 'cr', k = 5) +
    s(mu_hat, bs = 'cr', k = 5),
  family = gaussian(),
  data = d,
  method = 'fREML',
  knots = list(doy = c(0.5, 366.5)),
  drop.unused.levels = TRUE,
  discrete = TRUE,
  samfrac = 0.001,
  nthreads = future::availableCores(logical = FALSE) - 2,
  control = gam.control(trace = TRUE))
```

\normalsize

\doublespacing

# Results

output of `draw()`

hex plots of mean and var vs common metrics:

* temperature
* seasonal temperature range
* precip
* seasonal temperature range
* spp diversity
* gross primary productivity
* hfi
* max animal weight
* https://silvis.forest.wisc.edu/globalwui/
* https://silvis.forest.wisc.edu/data/dhis/: NDVI16 (Normalized Difference Vegetation Index), EVI16 (Enhanced Vegetation Index), FPAR8 (Fraction absorbed Photosynthetically Active Radiation), LAI8 (Leaf Area Index), GPP8 (Gross Primary Productivity)
* https://silvis.forest.wisc.edu/data/eurasia-trends/

# Discussion

* green wave surfing: change in the mean, failure to surf: variance
* migration routes are key habitat (Ortega et al., in prep), especially in stochastic environments
* @sawyer_migratory_2019
* https://www.sciencedirect.com/science/article/pii/S0960982220308484
* https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.15169
* industrial development affects animals' ability to respond to both predictable and unpredictable change (https://www.nature.com/articles/s41559-022-01887-9)
* effects of sudden deep snow in red desert on proghorn (https://www.sciencedirect.com/science/article/abs/pii/S0960982225002957)
* https://www.nature.com/articles/s41467-023-37750-z
* https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.4238
* see data for snow and pronghorn paper (Aikens et al. 2025)
* detecting regime shifts [in lakes: @bjorndahl_abrupt_2022]

* smoothness of mu impacts smoothness and size of s2: depends on animals' ability to respond to, learn, and predict E(R)

## Why not use machine learning? / Advantages of not using ML

arguably, GAMs can be considered as a form of machine learning, but we are referring to neural networks and other forms of black-box ML.

\clearpage

# Useful references

* `{khroma}`: @frerebeau_khroma_2024
* @keith_predicting_2008
* @chevin_adaptation_2010
* @rickbeil_plasticity_2019
* @mueller_search_2008
* @pettorelli_using_2005
* @keys_machine-learning_2021
* @nilsen_can_2005
* @stephens_optimal_1982
* @merkle_large_2016
* @tian_evaluating_2015
* @huang_commentary_2021
* @fan_global_2016
* @wang_stochastic_2019
* @pease_ecological_2024
* @xu_plasticity_2021
* @mezzini_how_2025

\clearpage

# References